---
title: "Correlation tests with R"
output: html_notebook
---

# PREVIOUS STEPS

Here is for installing missing packages and load them in the working session.

```{r}
pkgs2use<- c('haven','tidyverse','car','here','boot')

inst.load.pkg <- function(pkg){ 
  if (!require(pkg, character.only = TRUE)) {
      install.packages(pkg, dependencies = TRUE)
      library(pkg, character.only = TRUE)
  }
}


invisible(lapply(pkgs2use,inst.load.pkg))
rm(pkgs2use,inst.load.pkg)
```

# PEARSON'S CORRELATION TESTS

### Loading dataset

```{r}
dataEx1 <- haven::read_sav(here("MATERIALS","Practice","DataSets", "ExCor1.sav"))
```

```{r}
ggplot(dataEx1,aes(x=PANAS.NA,y=CFQ)) +
  geom_point() +
  geom_smooth()
```

### Correlation test

```{r}
with(dataEx1,cor.test(PANAS.NA,CFQ))
```


# SPEARMAN'S CORRELATION TESTS

### Loading dataset

```{r}
dataEx2 <- haven::read_sav(here("MATERIALS","Practice","DataSets", "ExCor2.sav"))
```

```{r}
ggplot(dataEx2,aes(x=qualitySN,y=riskSE)) +
  geom_point() +
  geom_smooth()
```

### Correlation test

```{r}
with(dataEx2,cor.test(qualitySN,riskSE,method='spearman'))
```

### Confidence interval for true correlation (Bootstrap)

```{r warning=FALSE,message=FALSE}
extract.rho <- function(df,i){
  cor.test(df[i,'x'],df[i,'y'],method="spearman")$estimate
}

df <- data.frame(x=dataEx2$qualitySN,y=dataEx2$riskSE)

corboot <- boot(df,extract.rho, R=5000)

boot.ci(corboot, type = c("bca"))

```


# KENDALL'S CORRELATION TESTS

### Loading dataset

```{r}
dataEx3 <- haven::read_sav(here("MATERIALS","Practice","DataSets", "ExCor3.sav"))
```

```{r}
dataEx3$education <- as_factor(dataEx3$education)
dataEx3$phase <- as_factor(dataEx3$phase)
df <- dataEx3 %>%
  group_by(education,phase) %>%
  summarise(counts = n(),.groups = 'drop') 

ggplot(df, aes(x = phase, y = counts)) +
  geom_bar(
    aes(color = education, fill = education),
    stat = "identity", position = position_dodge(0.8),
    width = 0.7) +
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  scale_fill_manual(values = c("#00AFBB", "#E7B800", "#FC4E07")) + theme_bw()
```

### Correlation test

```{r}
with(dataEx3,cor.test(as.numeric(phase),as.numeric(education),method='kendall'))
```

### Confidence interval for true correlation (Bootstrap)

```{r warning=FALSE,message=FALSE}
extract.tau <- function(df,i){
  cor.test(df[i,'x'],df[i,'y'],method="kendall")$estimate
}

df <- data.frame(x=as.numeric(dataEx3$phase),y=as.numeric(dataEx3$education))

corboot <- boot(df,extract.tau, R=5000)

boot.ci(corboot, type = c("bca"))

```

