---
title: "Mediation and Moderation Models with R"
output: html_notebook
---

# PREVIOUS STEPS

Before sourcing file **process.R**, you need to download it at the folder in which you've saved the Rmd file. The R script is available at [this link](https://www.processmacro.org/download.html).

```{r}
pkgs2use<- c('haven','tidyverse','here','DiagrammeR','interactions')

inst.load.pkg <- function(pkg){ 
  if (!require(pkg, character.only = TRUE)) {
      install.packages(pkg, dependencies = TRUE)
      library(pkg, character.only = TRUE)
  }
}


invisible(lapply(pkgs2use,inst.load.pkg))
rm(pkgs2use,inst.load.pkg)

source(here("MATERIALS","Practice","process.R"))
```

# SIMPLE MEDIATION MODELS WITH R

### Import dataset

```{r}
StressDF <- read_sav(here("MATERIALS","Practice","DataSets", "estress.sav"))
```

### Fit model

```{r message=FALSE,warning=FALSE}
results2 <- process(StressDF,y='withdraw',x='estress',m='affect',total=1,model=4,progress=0,save=3)
```

### Plot fitted model

```{r}
resCOEFS <- round(results[[2]],2) # This results[[2]] are availabe because we specified save=3!!!

starlabs <- c(paste("c´=",resCOEFS[12,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[12,4]<.001,'***',ifelse(resCOEFS[12,4]<.01,'**',ifelse(resCOEFS[12,4]<.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("a<FONT POINT-SIZE='6'><SUB>1</SUB></FONT>=",resCOEFS[3,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[3,4]<.001,'***',ifelse(resCOEFS[3,4]<.01,'**',ifelse(resCOEFS[3,4]<.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("b<FONT POINT-SIZE='6'><SUB>1</SUB></FONT>=",resCOEFS[7,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[7,4]<.001,'***',ifelse(resCOEFS[7,4]<.01,'**',ifelse(resCOEFS[7,4]<.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("a<FONT POINT-SIZE='6'><SUB>2</SUB></FONT>=",resCOEFS[6,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[6,4]<.001,'***',ifelse(resCOEFS[6,4]<.01,'**',ifelse(resCOEFS[6,4]<.05,'*',' '))),"</SUP></FONT>",sep='')
              )
grViz(paste("
      digraph SEM {
      
      graph [layout = neato,
      overlap = true,
      outputorder = edgesfirst
      title='TF']
      
      node [shape = rectangle,fontsize=10]
      
      
      a [pos = '-1,0!', label = 'STRESS']
      b [pos = '1,1!', label = 'AFFECT']
      c [pos = '3,0!', label = 'WITHDRAWAL']

      node [shape=plaintext,fontsize=9.5]
      
      ac [pos='1,.1!', label = <",starlabs[1],">]
      ab [pos='-0.2,.6!', label = <",starlabs[2],">]
      bc [pos='2.2,.6!', label = <",starlabs[3],">]
      
      a->b
      a->c
      b->c
}
",sep='')
)
```


# MULTIPLE MEDIATION MODELS WITH R

### Import dataset

```{r}
pmiDF <- read_sav(here("MATERIALS","Practice","DataSets", "pmi.sav"))
```

### Fit model

```{r}
results <- process(pmiDF,y='reaction',x='cond',m=c('pmi','import'),total=1,model=4,save=3,progress=0)
```

### Plot fitted model

```{r}
resCOEFS <- round(results[[2]],2) # This results[[2]] are available because we specified save=3!!!

starlabs <- c(paste("c´=",resCOEFS[9,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[9,4]<.001,'***',ifelse(resCOEFS[9,4]<.01,'**',ifelse(resCOEFS[9,4]<=.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("a<FONT POINT-SIZE='6'><SUB>1</SUB></FONT>=",resCOEFS[6,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[6,4]<.001,'***',ifelse(resCOEFS[6,4]<.01,'**',ifelse(resCOEFS[6,4]<=.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("b<FONT POINT-SIZE='6'><SUB>1</SUB></FONT>=",resCOEFS[11,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[11,4]<.001,'***',ifelse(resCOEFS[11,4]<.01,'**',ifelse(resCOEFS[11,4]<=.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("a<FONT POINT-SIZE='6'><SUB>2</SUB></FONT>=",resCOEFS[3,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[3,4]<.001,'***',ifelse(resCOEFS[3,4]<.01,'**',ifelse(resCOEFS[3,4]<=.05,'*',' '))),"</SUP></FONT>",sep=''),
              paste("b<FONT POINT-SIZE='6'><SUB>2</SUB></FONT>=",resCOEFS[10,1],"<FONT POINT-SIZE='6'><SUP>",ifelse(resCOEFS[10,4]<.001,'***',ifelse(resCOEFS[10,4]<.01,'**',ifelse(resCOEFS[10,4]<=.05,'*',' '))),"</SUP></FONT>",sep='')              
              )
grViz(paste("
      digraph SEM {
      
      graph [layout = neato,
      overlap = true,
      outputorder = edgesfirst
      title='TF']
      
      node [shape = rectangle,fontsize=10]
      
      
      a [pos = '-1,0!', label = 'CONDITION']
      b [pos = '1,1!', label = 'IMPORT']
      c [pos = '3,0!', label = 'REACTION']
      d [pos = '1,-1!', label = 'PMI']

      node [shape=plaintext,fontsize=9.5]
      
      ac [pos='1,.1!', label = <",starlabs[1],">]
      ab [pos='-0.2,.6!', label = <",starlabs[2],">]
      bc [pos='2.2,.6!', label = <",starlabs[3],">]
      ad [pos='0.3,-.4!', label = <",starlabs[4],">]
      dc [pos='1.8,-.4!', label = <",starlabs[5],">]
      
      a->b
      a->c
      b->c
      a->d
      d->c
}
",sep='')
)
```


# MODERATION MODELS WITH R

### Import dataset

```{r}
disasterDF <- read_sav(here("MATERIALS","Practice","DataSets", "disaster.sav"))
```

### Fit Model

```{r}
process(disasterDF,x='frame',y='justify',w='skeptic',model=1,jn=1)
```

### Plot moderation

```{r}
mod.moderation <- lm(justify~frame*skeptic,data=disasterDF)
interactions::johnson_neyman(model = mod.moderation, pred = frame,modx = skeptic)
```

